---
title: 'OHI `r format(Sys.Date(), "%Y")` - Tourism and Recreation '
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 1
    toc_float: yes
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '../../../workflow/templates/ohi_hdr.html' 
  pdf_document:
    toc: true
editor_options: 
  chunk_output_type: console
---


[REFERENCE RMD FILE](http://ohi-science.org/ohiprep_v2022/globalprep/tr/v2022/tr_data_prep.html)


# Summary
This document describes the steps for obtaining the data used to calculate the tourism and recreation goal for the 2022 global assessment.

The general calculation is:
tr = Ep * Sr 
and
Xtr = tr/90th quantile across regions

* Ep = Proportion of workforce directly employed in tourism
* Sr = (S-1)/5; Sustainability of tourism


## The following data are used:

* Tourism sustainability: World Economic Forum. The Travel & Tourism Development Index 2021 dataset (version 24 May 2022). 2022. [TTDI](https://www.weforum.org/reports/travel-and-tourism-development-index-2021/downloads-510eb47e12#report-nav)
* Proportion of workforce directly employed in tourism: obtained through the 2021 TTDI, however the WEF sources the employment data contained in the TTDI from the World Travel & Tourism Council ([WTTC](https://www.wttc.org/))
* Percent change in "Total contribution of Travel & Tourism to Employment", used to calculate proportion of workforce directly employed in tourism: scraped from PDFs on the [WTTC Economic Impact page](https://wttc.org/Research/Economic-Impact)
* Per capita GDP: (World Bank with gaps filled using CIA data), used to gapfill missing values in Tourism sustainability


# Updates from previous assessment

## Tourism sustainability
Data have been changed to the TTCI successor: the TTDI which includes a specific sustainability subindex. This new dataset only includes values for 2019 and 2021. 

## Tourism employment
2020 and 2021 direct contribution of tourism to employment values were calculated using 2019 values retrieved from WTTC in 2020 and percent change in total contribution of tourism to employment from WTTC economic reports.


**We were able to update the following data:**

* Tourism sustainability data from the WEF Travel and Tourism Development Index replaced data from its predecessor the TTCI and were updated to include 2021 and back-calculated 2019. (downloaded on 07/01/2022)

## Initial set-up code

```{r setup, message=FALSE, warning=FALSE, results="hide"}

# library(devtools)
# devtools::install_github("ohi-science/ohicore@dev") # dont worry about devtools
library(ohicore)
library(tidyverse)
library(stringr)
library(WDI)
library(here)
library(janitor)
library(plotly)
library(readxl)

version_year <- "2022"

source(paste0("http://ohi-science.org/ohiprep_v", version_year, "/workflow/R/common.R"))

## maximum year of wttc data:
year_max <- 2020

source(here(paste0("globalprep/tr/v", version_year, "/R/tr_fxns.R")))

```


# Ep: Proportion of workforce directly employed in tourism

We use "direct" employment data from the [World Travel & Tourism Council](http://www.wttc.org/). Up until the current assessment, we accessed the data directly from the WTTC — this is no longer an option so we are using the most recent of those data that we do have (2019), along with percent change in "Total contribution of Travel & Tourism to Employment" from the WTTC economic reports which they provide in pdf format for each country. We are assuming that percent change will be similar enough between these two measures in order for us to calculate direct contribution of tourism to employment for additional years (2020 & 2021).


### Calculating *direct* contribution of tourism to employment using WTTC percent change in contribution of tourism to employment

To make extracting data from the ~180 pdfs linked [here](https://wttc.org/Research/Economic-Impact) less time-consuming, the script sourced below will download all of them from the WTTC website, read them in as dataframes (which end up formatted non-traditionally, but consistently), and pull the desired percent change values into an output dataframe. That dataframe is then saved to the `intermediate` folder, and all the pdfs are deleted.

The percent change values are made proportions and used to calculate direct contribution of tourism to employment. This is done by multiplying the proportion change between 2019 and 2020 by the 2019 values we have to calculate 2020 values — the resulting 2020 values are multiplied by the proportion change between 2020 and 2021 to calculate 2021 values.


```{r}
# source script that gets percent change from WTTC pdfs - may take a couple minutes or so to do its thing
# source(here(paste0("globalprep/tr/v", version_year, "/R/wttc_pdf_scrape.R")))

tr_jobs_pct_v2021 <- read_csv(here("globalprep/tr/v2021/output/tr_jobs_pct_tourism.csv")) %>% 
  filter(year == 2019 & !is.na(Ep))

# read in csv created by above script and add rgn_ids to it
tr_jobs_pct_change <- read_csv(here(paste0("globalprep/tr/v", version_year,
                                           "/intermediate/tr_jobs_pct_change.csv"))) %>% 
  name_2_rgn(fld_name = "country") %>% 
  mutate(prop_change_2020 = pct_change_2020 / 100,
         prop_change_2021 = pct_change_2021 / 100) %>% 
  select(-c(pct_change_2020, pct_change_2021, rgn_name))

## get weighted mean values for duplicated regions
weight_data <- tribble(~country, ~rgn_id, ~population, # estimates from UN (http://data.un.org/Data.aspx?d=POP&f=tableCode%3a1)
                       "Guadeloupe", 140, 412682,
                       "Martinique", 140, 359821,
                       "Puerto Rico", 116, 3285874, 
                       "US Virgin Islands", 116, 106669,
                       "China", 209, 1412360000,
                       "Hong Kong SAR", 209, 7413100)

tr_jobs_pct_change <- tr_jobs_pct_change %>%
  arrange(country) %>%
  left_join(weight_data, by = c("country", "rgn_id")) %>%
  mutate(population = ifelse(is.na(population), 1, population)) %>%
  group_by(rgn_id) %>%
  summarize(across(c(prop_change_2020, prop_change_2021), ~ weighted.mean(.x, population)))


# calculate 2020 Ep estimates using 2019 values and proportion change, and then repeat for 2021 values using the 2020 estimates
tr_jobs_pct_tour <- tr_jobs_pct_v2021 %>% 
  left_join(tr_jobs_pct_change) %>% 
  pivot_wider(names_from = year, values_from = Ep) %>% 
  rename("Ep_2019" = "2019") %>% 
  mutate("Ep_2020" = Ep_2019 * (1 + prop_change_2020)) %>% 
  mutate("Ep_2021" = Ep_2020 * (1 + prop_change_2021)) %>% 
  select(-prop_change_2020, -prop_change_2021, "2019" = Ep_2019, "2020" = Ep_2020, "2021" = Ep_2021) %>% 
  pivot_longer(cols = "2019":"2021", names_to = "year", values_to = "Ep")


```


### Gapfilling using UN georegions

```{r}
## gapfill missing data using UN georegion data:
georegions <- georegions
georegion_labels <- georegion_labels

tr_jobs_pct_gf <- tr_jobs_pct_tour %>%
  left_join(georegions, by = 'rgn_id') %>%
  left_join(georegion_labels, by = 'rgn_id') %>%
  select(-r0)

# Calculate two different gapfill columns using r2 and r1
tr_jobs_pct_gf <- tr_jobs_pct_gf %>%
  group_by(year, r2) %>%
  mutate(Ep_pred_r2 = mean(Ep, na.rm=TRUE)) %>%
  ungroup() %>%
  group_by(year, r1) %>%
  mutate(Ep_pred_r1 = mean(Ep, na.rm=TRUE)) %>%
  ungroup()

# first gapfill with r2, if no value available use r1; create column indicating whether value was gapfilled and if so, by what method.
tr_jobs_pct_gf <- tr_jobs_pct_gf %>%
  mutate(Ep_all = ifelse(is.na(Ep), Ep_pred_r2, Ep)) %>%
  mutate(Ep_all = ifelse(is.na(Ep_all), Ep_pred_r1, Ep_all)) %>% 
  mutate(gapfilled = case_when(is.na(Ep) & !is.na(Ep_all) ~ "gapfilled")) %>%
  mutate(method = case_when(is.na(Ep) & !is.na(Ep_pred_r2) ~ "UN georegion (r2)",
                            is.na(Ep) & is.na(Ep_pred_r2) & !is.na(Ep_pred_r1) ~ "UN georegion (r1)"))

```

### Removing low population / uninhabited regions

```{r}

### After gap-filling, make sure low/uninhabited regions are NA
# Create df for unpopulated/low populated regions
low_pop()
low_pop <- low_pop %>%
  filter(est_population < 3000 | is.na(est_population)) %>%  #filter out regions that have populations > 3000 and keep NA values 
  rename(rgn_label = rgn_nam)

summary(tr_jobs_pct_gf)
# v2020 371 NAs
# v2022 114 NAs

# make sure all the NAs are uninhabited regions
tr_jobs_pct_nas <- tr_jobs_pct_gf %>% 
  filter(is.na(Ep_all)) %>% 
  select(rgn_id, year, r1_label, r2_label, rgn_label) %>% 
  left_join(low_pop, by = c("rgn_id", "rgn_label"))

tr_jobs_pct_nas %>% 
  filter(Inhabited == 0 & !is.na(est_population)) %>% 
  nrow() # 0 ✓

max(tr_jobs_pct_nas$est_population, na.rm=TRUE) < 3000 # should be true

# make sure all the uninhabited regions are NA (along with gapfill and method if they were gapfilled above)
tr_jobs_pct_gf <- tr_jobs_pct_gf %>% 
  mutate(Ep_all = ifelse(rgn_id %in% low_pop$rgn_id, NA, Ep_all)) %>% 
  mutate(gapfilled = ifelse(is.na(Ep_all), NA, gapfilled)) %>% 
  mutate(method = ifelse(is.na(Ep_all), NA, method))


# check NAs once more 
summary(tr_jobs_pct_gf)
# v2019: Adding the low pop df identifies 13 additional regions that should be NA instead of gapfilled, taking the total number of NAs in the data set from 245 to 700
# v2020: Adding the low pop df takes the total number of NAs in the data set from 371 to 832
# v2022: Adding the low pop df takes the total number of NAs in the data set from 14 to 40

```

### Bind data from previous years and write output files

```{r}
# read in data from previous years
tr_jobs_prev_years <- read_csv(here("globalprep/tr/v2020/output/tr_jobs_pct_tourism.csv"))
tr_jobs_prev_years_gf <- read_csv(here("globalprep/tr/v2020/output/tr_jobs_pct_tourism_gf.csv"))

tr_jobs_prev_combined <- tr_jobs_prev_years %>% 
  left_join(tr_jobs_prev_years_gf) %>% 
  filter(year < 2019 & year >= 2008)

tr_jobs_pct_gf <- tr_jobs_pct_gf %>%
  select(rgn_id, Ep = Ep_all, year, gapfilled, method) %>% 
  rbind.data.frame(tr_jobs_prev_combined)

tr_jobs_pct_tourism_gf <- tr_jobs_pct_gf %>%
  select(rgn_id, year, gapfilled, method)

# save gapfill info
write_csv(tr_jobs_pct_tourism_gf, here(paste0("globalprep/tr/v", version_year, "/output/tr_jobs_pct_tourism_gf.csv")))

tr_jobs_pct_tourism <- tr_jobs_pct_gf %>%
  select(rgn_id, year, Ep) 

# Save gap-filled data
write_csv(tr_jobs_pct_tourism, here(paste0("globalprep/tr/v", version_year, "/output/tr_jobs_pct_tourism.csv")))
```

### Look at changes in recent years

We would expect for tourism jobs to decrease across the board from 2019 and 2020 given the pandemic, and likely see a rebound to some extent between 2020 and 2021 — let's make sure that's reflected in our results.

```{r}

tr_jobs_pct_compare <- tr_jobs_pct_tourism %>% 
  filter(year >= 2019) %>%
  pivot_wider(names_from = year, values_from = Ep)

# compare 2019 and 2020
plot(tr_jobs_pct_compare$"2019", tr_jobs_pct_compare$"2020",
     xlab = "2019 Employment Proportion", ylab = "2020 Employment Proportion")
abline(0, 1)

# compare 2020 and 2021
plot(tr_jobs_pct_compare$"2020", tr_jobs_pct_compare$"2021",
     xlab = "2020 Employment Proportion", ylab = "2021 Employment Proportion")
abline(0, 1)

```

Everything looks reasonable.


# Ts: Tourism sustainability

These data are from the World Economic Forum's "Travel and Tourism Development Index" (https://www.weforum.org/reports/travel-and-tourism-development-index-2021/downloads-510eb47e12) See mazu: _raw_data/WEF-Economics/ for more details and the raw data.

The TTDI was formerly the TTCI which was a similar index, but unfortunately not comparable. The TTDI only extends back to 2019. 

These data are gapfilled using gdppcppp and UN georegion information (see next section for obtaining and preparing these data).

```{r WEF processing, eval=FALSE}
# update to latest file name
ttdi_file <- "WEF_TTDI_2021_data_for_download.xlsx"

ttdi_raw <- read_excel(paste0(dir_M, "/git-annex/globalprep/_raw_data/WEF-Economics/d", version_year, "/", ttdi_file),
                 skip = 2) 

# move up column names from first row while keeping the full country names as columns too
names(ttdi_raw)[1:9] <- as.character(ttdi_raw[1, 1:9])

# filtering for sustainability scores, selecting needed columns, and pivoting to tidy format
ttdi <- ttdi_raw %>%
  filter(Title == "T&T Sustainability subindex, 1-7 (best)",
         Attribute == "Score") %>% 
  select(year = Edition, Albania:Zambia) %>% 
  # currently Zambia is the last country column
  pivot_longer(cols = Albania:Zambia, names_to = "country",
               values_to = "score") %>% 
  mutate(score = as.numeric(score))


# Changing names that are not recognized by ohicore
ttdi <- ttdi %>%
    mutate(country = ifelse(str_detect(country, "Ivoire"), "Ivory Coast", country))
  
  
ttdi_rgn <- name_2_rgn(df_in = ttdi, 
                       fld_name='country')

## Duplicated regions weighted mean
weight_data <- data.frame(country = c("China", "Hong Kong SAR"),
                          # pop values from World Bank 2021 estimates - updated v2022
                          population = c(1412360000, 7413100))


ttdi_rgn <- ttdi_rgn %>%
  arrange(country) %>%
  left_join(weight_data, by = "country") %>%
  mutate(population = ifelse(is.na(population), 1, population)) %>%
  group_by(rgn_id, rgn_name, year) %>%
  summarize(score = weighted.mean(score, population)) %>%
  select(year, rgn_id, rgn_name, score)

# compare with old dataframe to make sure only the duplicated region scores changed 

head(ttdi_rgn, 10)

### Save TTDI data file
write.csv(ttdi_rgn, here(paste0("globalprep/tr/v", version_year, "/intermediate/wef_ttdi.csv")), row.names = FALSE)

```

## Preparing the gdppcppp data:
These data are used to gapfill missing values in tourism sustainability.  Most of the data are from the World Bank, but CIA data fill some gaps (CIA data is available for only the most recent year).

The Artisanal Opportunities goal uses gdppcppp data, so we will get the data that was processed for that goal.


```{r worldbank, eval=FALSE}
wb <- read.csv(here(paste0("globalprep/ao/v", version_year, "/intermediate/gdppcppp_ohi.csv"))) %>%
  dplyr::select(rgn_id, year, value)

```

CIA data are used to fill in missing gaps in the gdppcppp data (https://www.cia.gov/the-world-factbook/field/real-gdp-per-capita/country-comparison)

Downloaded: 07/05/2022

See README on the raw folder for instructions on how to download this data. 

The following code is used to prepare these data for OHI:

```{r cia gdp, eval=FALSE}

cia_gdp <- read.csv(here(paste0("globalprep/tr/v", version_year, "/raw/cia_gdp_pc_ppp.csv"))) %>% 
  # remove dollar signs and commas and convert to numeric
  mutate(value = as.numeric(gsub("[$,]", "", value))) %>% 
  select(name, value) %>% 
  rename(country = name, pcgdp_cia = value)

 ## Data reported in a lower resolution than OHI regions
splits <- data.frame(country = "Saint Helena, Ascension, and Tristan da Cunha", 
                     country2 = c("Saint Helena", "Ascension","Tristan da Cunha"))

cia_gdp <- cia_gdp %>%
  left_join(splits, by='country') %>%
  mutate(country2 = ifelse(is.na(country2), country, country2)) %>%
  select(country = country2, pcgdp_cia)

cia_gdp_rgn <- name_2_rgn(df_in = cia_gdp, 
                       fld_name='country')

### Duplicated regions: Collapse regions after weighting by population (regions we include as a single region) - 

population_weights <- data.frame(country = c("Virgin Islands", "Puerto Rico",
                                             "China", "Hong Kong", "Macau",
                                             "Guam", "Northern Mariana Islands"),
                                 # from world bank - updated v2022
                                 population = c(105870, 3263584, 1412360000,
                                                7413100, 658391, 170184, 57910))

cia_gdp_rgn <- cia_gdp_rgn %>%
  left_join(population_weights, by="country") %>%
  mutate(population = ifelse(is.na(population), 1, population)) %>%
  group_by(rgn_id) %>%
  summarize(pcgdp_cia = weighted.mean(pcgdp_cia, population)) %>%
  ungroup() %>%
  filter(rgn_id <= 250) %>%
  select(rgn_id, pcgdp_cia)

write.csv(cia_gdp_rgn, here(paste0("globalprep/tr/v", version_year, "/intermediate/wb_rgn_cia_GDPPCPPP.csv")), row.names=FALSE)

```

The following code combines the two gdp datasets and gapfills missing regions using UN georegions.

If there is no World Bank gdppcppp data (pcgdp), the CIA data is used (pcgdp_cia).  The pcgdp2 variable includes both the World Bank and CIA data (with CIA data only used if there is not World Bank data).  The remaining data are estimated using UN geopolitical regions.  Ideally, the mean gdppcppp value is calculated at the r2 scale (gdp_pred_r2) using regions within each class with gdppcppp data.  If there were not enough regions with data at the r2 scale, the average at the r1 scale was used (gdp_pred_r1). The gdp_all variable combines all estimates using the following heirarchy:  World Bank -> CIA -> estimated using mean from r2 UN geopolitical regions -> estimated using mean from r1 UN geopolitical regions.    

```{r gapfill gdp, eval=FALSE}

### world bank gdp data
gdppcppp <- wb %>%
  select(rgn_id, year, pcgdp = value)

### cia gdp data
gdppcppp2 <- read.csv(here(paste0("globalprep/tr/v", version_year, "/intermediate/wb_rgn_cia_GDPPCPPP.csv")))


### Use WB data, but if missing, use pcgdp_cia.
### combine with UN georegion data
years <- data.frame(year = min(gdppcppp$year):max(gdppcppp$year))

georegions <- ohicore::georegions

regions <- georegions %>%
  left_join(georegion_labels, by = 'rgn_id')

gdp_raw <- merge(years, regions, by=NULL) %>%
   left_join(gdppcppp, by = c('rgn_id', 'year')) %>%
  left_join(gdppcppp2, by = c("rgn_id")) 

## quick compare to make sure the CIA and World Bank data are compatible
plot(gdp_raw$pcgdp[gdp_raw$year==2021], gdp_raw$pcgdp_cia[gdp_raw$year==2021])
abline(0,1, col="red")
# a few minor outliers but overall looks good

gdp_raw <- gdp_raw %>%
  mutate(pcgdp2 = ifelse(is.na(pcgdp), pcgdp_cia, pcgdp))

## Calculating the means across different geopolitical levels (e.g. r2, r1)
gdp_raw <- gdp_raw %>%
  group_by(r2, year) %>%
  mutate(gdp_pred_r2 = mean(pcgdp2, na.rm=TRUE)) %>%
  ungroup() %>%
  group_by(r1, year) %>%
  mutate(gdp_pred_r1 = mean(pcgdp2, na.rm=TRUE)) %>%
  ungroup() 

gdp_raw_gf <- gdp_raw %>%
  mutate(gdp_all = ifelse(is.na(pcgdp2), gdp_pred_r2, pcgdp2)) %>%
  mutate(gdp_all = ifelse(is.na(gdp_all), gdp_pred_r1, gdp_all)) %>%
  mutate(gapfilled = ifelse(is.na(pcgdp2) & !is.na(gdp_all), "gapfilled", NA)) %>%
  mutate(method = ifelse(is.na(pcgdp2) & !is.na(gdp_pred_r2), "UN georegion (r2)", NA)) %>%
  mutate(method = ifelse(is.na(pcgdp2) & is.na(gdp_pred_r2) & !is.na(gdp_pred_r1), "UN georegion (r1)", method)) 

write_csv(gdp_raw_gf, here(paste0("globalprep/tr/v", version_year, "/intermediate/gdp_raw_gf.csv")))

gdp_data_gf <- gdp_raw_gf %>%
  select(rgn_id, year, gapfilled, method) 

write_csv(gdp_data_gf, here(paste0("globalprep/tr/v", version_year, "/intermediate/gdp_gf.csv")))

gdp_data <- gdp_raw_gf %>%
  select(rgn_id, year, pcgdp = gdp_all)

write_csv(gdp_data, here(paste0("globalprep/tr/v", version_year, "/intermediate/gdp.csv")))

```


The final step is gapfilling the Sustainability data using a linear model with gdppcppp and UN geopolitical regions as predictor variables.  

```{r, eval=FALSE}

sust <- read.csv(here(paste0("globalprep/tr/v", version_year, "/intermediate/wef_ttdi.csv")), stringsAsFactors = FALSE)

### don't need to gapfill data without tourism data:
## Most recent tourism data is 2019.  

ep_gf <- read.csv(here(paste0("globalprep/tr/v", version_year, "/output/tr_jobs_pct_tourism.csv"))) %>%
  # filter(year == 2021) %>%
  select(rgn_id, Ep, year) %>%
  filter(!is.na(Ep))

# gdp dataframe prepared above (World Bank, CIA, and gapfilled gdp data)
gdp_raw_gf <- read.csv(here(paste0("globalprep/tr/v", version_year, "/intermediate/gdp_raw_gf.csv")), stringsAsFactors = FALSE) %>% 
  # filter(year == 2021) %>%
  select(rgn_id, r0_label, r1_label, r2_label, rgn_label,
         pcgdp, pcgdp_cia, pcgdp2, gdp_all, year) 

tr_sust <- gdp_raw_gf %>%
           left_join(sust, by = c("rgn_id", "year")) %>%
          left_join(ep_gf, by = c("rgn_id", "year")) %>%
          rename(S_score = score) %>%
          filter(rgn_id != 213)

### Add gapfill flag variable 
## Reminder:
## pcgdp2: includes both the World Bank and CIA data (with CIA data only used if there is not World Bank data)
## Ep: Proportion of workforce directly employed in tourism
## S_score: tourism sustainability score

tr_sust_gf <- tr_sust %>%
  mutate(gapfilled = ifelse(is.na(S_score) & !is.na(Ep), "gapfilled", NA)) %>%
  mutate(method = ifelse(is.na(S_score) & !is.na(Ep) & is.na(pcgdp2), "lm georegion + gdppcppp, with est. gdppcppp", NA)) %>%
  mutate(method = ifelse(is.na(S_score) & !is.na(Ep) & !is.na(pcgdp2), "lm georegion + gdppcppp", method)) %>%
  select(rgn_id, gapfilled, method, year)

write.csv(tr_sust_gf, here(paste0("globalprep/tr/v", version_year, "/output/tr_sustainability_gf.csv")), row.names=FALSE)

```
  

### Gapfilling
Linear models using gdppcppp and UN geopolitical regions as predictor variables. However if there is no gdppc data we estimate the gdppc using the UN georegions and then used in the linear model to gapfill the sustainability score.

```{r, eval=FALSE}

### Gapfill S using r1 and/or r2 regional data and PPP-adjusted per-capita GDP
### Looked at models with a year variable, but wasn't significant and decided to exclude

mod3 <- lm(S_score ~ as.factor(r2_label) + gdp_all, data=tr_sust, na.action = na.exclude)
summary(mod3)
anova(mod3)

mod4 <- lm(S_score ~ as.factor(r1_label) + gdp_all, data=tr_sust, na.action = na.exclude)
summary(mod4)
anova(mod4)

plot(predict(mod3), tr_sust$S_score)
abline(0,1)
plot(predict(mod4), tr_sust$S_score)
abline(0,1)


## Estimate missing data and gapfill
# Some of the r1 levels do not have data and consequently causes a fail. This chunk of code drops these levels so an NA is returned

# Select only r2 column
new_data <- tr_sust %>% 
  dplyr::select(r2_label, gdp_all)

unique(tr_sust$r2_label)

r2_w_data <- unique(tr_sust$r2_label[!is.na(tr_sust$S_score)])
  
new_data_r2 <- new_data %>%
  mutate(r2_label = ifelse(r2_label %in% r2_w_data, r2_label, NA))

# Predict sustainability scores using linear model 3 (using r2 data)
tr_sust <- tr_sust %>% 
  dplyr::mutate(S_score_pred_r2 = predict(mod3, newdata = new_data_r2))


# Select only r1 column
new_data <- tr_sust %>% 
  dplyr::select(r1_label, gdp_all)

unique(tr_sust$r1_label)

r1_w_data <- unique(tr_sust$r1_label[!is.na(tr_sust$S_score)])

new_data_r1 <- new_data %>%
  mutate(r1_label = ifelse(r1_label %in% r1_w_data, r1_label, NA))

# Predict sustainability scores using linear model 4 (using r1 data)
tr_sust <- tr_sust %>% 
  dplyr::mutate(S_score_pred_r1 = predict(mod4, newdata = new_data_r1))



## some are missing the r1 predictions, but none of these have Ep scores, so not relevant
View(filter(tr_sust, is.na(S_score_pred_r1)))

tr_sust <- tr_sust %>%
  mutate(S_score_2 = ifelse(is.na(S_score), S_score_pred_r2, S_score)) %>%
  mutate(S_score_2 = ifelse(is.na(S_score_2), S_score_pred_r1, S_score_2)) %>%
  filter(year %in% c(2019, 2021)) %>%
  select(rgn_id, year, S_score=S_score_2)

summary(tr_sust)

write_csv(tr_sust, here(paste0("globalprep/tr/v", version_year, "/output/tr_sustainability.csv")))
```


## Compare with previous year of data 

```{r, eval=FALSE}
tr_sust <- read_csv(here(paste0("globalprep/tr/v", version_year, "/output/tr_sustainability.csv")))

prev_year <- (as.numeric(version_year) - 1) %>% 
  as.character()

compare <- tr_sust %>% 
  pivot_wider(names_from = year, values_from = S_score)

# current vs previous year of data
plot(compare$"2021", compare$"2019")
abline(0, 1, col="red")
# looks good

```


# Tw: Travel warnings

 - Travel warnings were deleted from the v2020 assessment. 

